---
title: "Protein structure-based homology and machine learning"
subtitle: "Scripts to reproduce figures and tables in manuscript GIGA-D-18-00288"
author: "Brendan R.E. Ansell"
date: "10/16/2018"
output: 
  html_notebook:
    #theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
  #github_document: default
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, tidy = TRUE)


library(tidyverse); library(ggplot2); 
library(stringr); library(forcats)
library(caret); library(UpSetR)
library(rpart); library(pROC)


select <- dplyr::select

```

#Data import 

Read in data sets for BLAST annotation, PFAM annotation and I-TASSER output metrics
```{r, eval=TRUE}
GeneDesc <- read_tsv("../data/GiardiaDB-39_GintestinalisAssemblageA_AnnotatedProteins.description",
                     col_names = FALSE, col_types = cols())

names(GeneDesc) <- c('GeneID','Description')

metrics_df <- read.delim("../data/metrics_clean.txt",sep="\t", stringsAsFactors = FALSE, quote="", header=TRUE)

mismatch_BLAST <- read_tsv("../data/mismatches_BLAST.tsv", col_types = cols())

PDB_pfam <- read_tsv("../data/hmmer_pdb_all.txt", col_types = cols())  #accessed in Oct 2017

PDB_pfam <- PDB_pfam %>% mutate(code_chain=paste0(PDB_ID,CHAIN_ID)) %>%  
            separate(PFAM_ACC, into=c('PFAMpref','PFAMsuff')) 

###Giardia PFAM domains

#GDB_IPS <- read_tsv("http://giardiadb.org/common/downloads/release-39/GintestinalisAssemblageAWB/txt/GiardiaDB-39_GintestinalisAssemblageAWB_InterproDomains.txt", col_names = FALSE) 

GDB_IPS <- read_tsv("../data/GiardiaDB-39_GintestinalisAssemblageAWB_InterproDomains.txt", 
                    col_types = cols(), col_names = FALSE)
names(GDB_IPS) <- c('GeneID','source','InterproID','Desc','GDBpfam_start','GDBpfam_end','eVal')

GDB_pfam <- GDB_IPS %>% filter(str_detect(InterproID,'^PF')) %>%
  dplyr::rename(PFAMpref=InterproID) %>% distinct()

PFAM_descriptionTable <- read_tsv("../data/PFAM_descriptionMapping.tsv",col_names = TRUE, col_types = cols())

```

##Compute additional metrics

Metrics derived from I-TASSER output: sd of secondary structure predictions, and query:reference length ratio.
```{r, eval=TRUE}

metrics_all <- metrics_df %>% 
  mutate(lenRatio=seq.ssLen/PDB_AA) %>% 
  mutate(Hprop=nHrc/seq.ssLen,
                      Eprop=nErc/seq.ssLen, 
                      Cprop=nCrc/seq.ssLen) %>%  
  rowwise() %>% mutate(SS_sd = sd(c(Hprop, Eprop, Cprop))) 

```



#Check positive controls
##SFigure 1

Investigate discordant reference structure matches for *Giardia* peptides encoding solved structures

```{r}


mismatch_BLAST_status <- mismatch_BLAST %>% 
  mutate(matchStatus = factor(case_when(str_detect(status,'miss') ~ 'Matched_other',
                                   str_detect(status,'hit') ~ 'Matched_Gd'))) %>% 
  mutate(refSpecies = factor(case_when(str_detect(status,'hit') ~ 'Gd',
                                       str_detect(status,'non-Giardia') ~ 'other',
                                       TRUE ~ 'Gd'))) %>% 
  select(GeneID, matchStatus,refSpecies,Ref_Query_AAlenRatio,bit_score)
  
mismatch_BLAST_status %>% 
  rename(`Reference species` = refSpecies,
         `Match status` = matchStatus,
         `AA length ratio` = Ref_Query_AAlenRatio,
         'Bit score' = bit_score) %>% 
  gather(key,value,-c(GeneID,`Match status`,`Reference species`)) %>% 
  mutate(xSep=paste0(`Match status`,'_vs_',`Reference species`)) %>% 
  ggplot(aes(x=xSep, y=value)) + 
  geom_boxplot(aes(col=`Match status`), alpha=0) + 
  geom_jitter(aes(col=`Match status`,shape=`Reference species`), width=0.1, size=2) +
  facet_wrap(~ key, scales="free")   +
  theme(axis.text.x = element_text(angle=45,hjust=1)) +
  xlab("")

ggsave("../results/SF1.pdf",width=8,height=4)

```


Test differences in BLAST bit scores, and query:reference AA length ratios
```{r}

mismatch_BLAST_forTest <- mismatch_BLAST_status %>% unite(match_species, c(matchStatus,refSpecies))

#Test differences in Ref_Query_AAlenRatio
aov_lenRat.res <- aov(Ref_Query_AAlenRatio ~ match_species, data = mismatch_BLAST_forTest )
summary(aov_lenRat.res)

#Multiple pair-wise comparisons
TukeyHSD(aov_lenRat.res)

#Test differences in bit_score
aov_bitScore.res <- aov(bit_score ~ match_species, data = mismatch_BLAST_forTest )
summary(aov_bitScore.res)

#Multiple pair-wise comparisons
TukeyHSD(aov_bitScore.res)



```


#PFAM code matching

Join PFAM annotation databases and identify query:reference pairs with at least 1 matching PFAM code.  
Annotation: PFAM codes assigned to Giardia query peptides: 'GDB_pfam'   
            PFAM codes assigned to peptides encoding PDB reference structures: 'PDB_pfam'.

```{r, eval=TRUE}

metrics_long <- metrics_all %>% 
  mutate(code_chain=toupper(codechain)) %>% 
  left_join(GDB_pfam,by="GeneID") %>% rename(GDB_PFAMpref=PFAMpref) %>% 
  left_join(PDB_pfam,by="code_chain") %>% rename(PDB_PFAMpref=PFAMpref) %>% 
  distinct() %>% rename(GDB_PFAM_desc = Desc, PDB_PFAM_desc = PFAM_desc)


```

Plot number of PFAM domains per query:reference pair
```{r}
nPFAM_Q_R <- metrics_long %>% 
  select(GeneID, GDB_PFAMpref, PDB_PFAMpref) %>% 
  distinct() %>% 
  group_by(GeneID) %>% 
  summarize_at(vars(GDB_PFAMpref, PDB_PFAMpref), funs( . %>% na.omit() %>% n_distinct )) %>% 
  rename(nPFAM_GDB = GDB_PFAMpref, nPFAM_PDB = PDB_PFAMpref )

nPFAM_Q_R %>% gather(key,value,-GeneID) %>% ggplot(aes(value)) +
  geom_bar(aes(fill=key), position="dodge", show.legend = TRUE) 
```


Calculate n. matching domains
```{r}

#Because all pairwise combinations of PFAM domains are listed, need only filter for exact match:
nMatch_count <- metrics_long %>% 
  select(GeneID,code,chain,code_chain,GDB_PFAMpref,PDB_PFAMpref) %>%
  distinct() %>% 
  group_by(GeneID) %>% 
  filter(GDB_PFAMpref==PDB_PFAMpref) %>% 
  summarize(n_Match = n()) 
  
nMatch <- metrics_all[ , "GeneID"] %>% 
  left_join(nMatch_count, by="GeneID") %>% 
  mutate(n_Match = ifelse(is.na(n_Match), 0 ,n_Match)) 

```

Summarize PFAM matches across annotation groups
```{r}

metrics_nMatch <- metrics_all %>% 
  left_join(nPFAM_Q_R, by="GeneID") %>% 
  left_join(nMatch,by="GeneID") %>% 
   mutate(n_Match=ifelse(is.na(n_Match), 0, n_Match)) %>% 
   mutate(matchStatus=ifelse(n_Match==0,'noMatch','exactMatch')) %>% 
   left_join(GeneDesc,by="GeneID") %>% select(GeneID,Description,everything()) %>% 
   mutate(descBin = ifelse(str_detect(Description,'hypothetical') | str_detect(Description,'Hypothetical'), 'Hyp','Annot')) 
```

```{r}
metrics_nMatch %>% count(descBin,matchStatus) 
```


##SFigure 2
Relative peptide length and number of matching PFAM domains.
```{r}
lenRatio_X_nMatch <- metrics_nMatch %>% 
  select(GeneID, Cov, GDB_AA, lenRatio, contains('nPFAM'),n_Match, descBin) %>% 
  mutate(`n_Match:nPFAM_PDB` = n_Match/nPFAM_PDB) %>% 
  na.omit() %>% 
  group_by(n_Match, nPFAM_PDB) %>% summarize(median_lenRatio = median(lenRatio), 
                                                mean_lenRatio = mean(lenRatio), 
                                                mean_GDB_AAlen = mean(GDB_AA), 
                                                median_GDB_AAlen = median(GDB_AA), 
                                             group_size = n()) %>% 
  arrange(n_Match, nPFAM_PDB, desc(group_size))


metrics_nMatch %>% 
  filter(nPFAM_PDB<=4) %>% #filter(n_Match==0) %>% 
  filter(nPFAM_PDB>0) %>% 
  ggplot() + 
  geom_boxplot(aes(x=n_Match , y=lenRatio, group=factor(n_Match)), alpha=0, width=0.5)+  
  geom_jitter(aes(x=n_Match , y=lenRatio, col=factor(nPFAM_PDB), 
                  group=factor(nPFAM_PDB)), 
                  position=position_jitterdodge(dodge.width = 0.85), alpha=0.5, cex=0.25) +
  geom_text(data=lenRatio_X_nMatch %>% 
              filter(n_Match<=4, nPFAM_PDB<=4), 
            aes(x=n_Match, y=3, label=group_size, group=nPFAM_PDB), size=3) +
  facet_wrap( ~ nPFAM_PDB) +
  theme(legend.position = "none") +
  ggtitle("N. matching PFAM terms vs query:reference AA length ratio",
          subtitle="Facet by n. PFAM terms in PDB reference") +
  xlab("N. matching PFAM domains") + ylab("Query:reference AA length ratio")

ggsave("../results/SF2.pdf",width=6,height=6)

```


##Figure 2a
Plot intersection of BLAST annotation status and PFAM code annotations.

```{r}
source("upset.R")

upSets <- metrics_nMatch %>% select(GeneID, descBin, nPFAM_GDB, nPFAM_PDB, n_Match) %>% 
  mutate_at(vars(-GeneID, -descBin) , funs(binary = ifelse(. > 0, 1, 0))) %>% 
  mutate(descCategory = ifelse(descBin=="Hyp",1,2)) %>%  
  distinct() %>% select(GeneID, descCategory, contains('binary')) %>% 
  rename(Query = nPFAM_GDB_binary, Reference = nPFAM_PDB_binary, Match = n_Match_binary)

upSets %>% count(descCategory, Query, Reference, Match)        

annotStatus <- function(row, min, max){
  newData <- (row["descCategory"] <= max) & (row["descCategory"] > min)}

upset(data.frame(upSets), 
      sets=c('Query','Reference','Match'),
      main.bar.color = 'black',
      queries = list(list(query = annotStatus, params = list(1,2),
                          color="light blue",active=TRUE)), point.size=3)

pdf("../results/F2a.pdf", width = 4,height=3, onefile = FALSE) #works..ish
#pdf("../results/F2a.pdf", width = 4.5,height=3.5, onefile = FALSE)
upset(data.frame(upSets), 
      sets=c('Query','Reference','Match'),
      main.bar.color = 'black',
      queries = list(list(query = annotStatus, params = list(1,2),
                          color="light blue",active=TRUE)), point.size=3)

dev.off()

```



##Figure 2b

PFAM term enrichment testing


```{r}
source("rowwise_fisher.R")

matchStatus_PFAMcounts <- metrics_long %>% 
  left_join(metrics_nMatch %>% select(GeneID, matchStatus), by="GeneID") %>% 
  dplyr::select(1,lenRatio, matchStatus, GDB_PFAMpref,PDB_PFAMpref) %>% 
  gather(key=DB,value=PFAM,-c(GeneID,matchStatus,lenRatio)) %>% 
  na.omit() %>% 
  filter(lenRatio >0.9 & lenRatio <1.1) %>% ungroup() %>% 
  count(matchStatus,DB,PFAM) %>% 
  spread(key=DB, value=n, fill = 0) 

#set up parameters for Fisher tables  
myStatus <- "exactMatch"
minFreq <- 3
Foreground <- 'PDB_PFAMpref'
background <- 'GDB_PFAMpref'

termSums <- matchStatus_PFAMcounts %>% 
  filter(get(Foreground) >= minFreq) %>%
  filter(matchStatus == myStatus) %>% 
  summarize(allForeground = sum(get(Foreground)),
            allBackground = sum(get(background)),
            allTerms = sum(get(Foreground),get(background))) %>% unlist()

Fisher_df <- matchStatus_PFAMcounts %>% 
  filter(matchStatus==myStatus) %>% 
  dplyr::rename(Foreground = PDB_PFAMpref, 
                BACKGROUND = GDB_PFAMpref) %>% 
  rowwise() %>% 
  mutate(sumTerm=sum(Foreground,BACKGROUND),
         allForeground=termSums[[1]],
         allBackground=termSums[[2]],
         allTerms=termSums[[3]]) %>% 
  mutate('a' = Foreground,
         'b' = sumTerm - Foreground,
         'c' = allForeground - Foreground,
         'd' = allTerms - allForeground - b) %>% 
  dplyr::select(PFAM,Foreground,BACKGROUND,everything()) %>% 
  ungroup() 


mydata <- Fisher_df
p <- t(apply(mydata %>% 
               dplyr::select(a,b,c,d) %>% 
               filter(a >= minFreq), 1, row_fisher))


results <- cbind(mydata %>% filter(a >= minFreq) %>% 
                   dplyr::select(1:7), p) %>% 
#1:7 as no background of repressed, lowly transcribed (<0.85 percentile; i.e., the equivalent of the nonDTG set), is included.
  arrange(p_val) %>% 
  mutate(adjP=p.adjust(.$p_val, method="BH" ))  #correction for multiple comparisons

#No significant enrichment
results %>% filter(p_val < 0.05) %>% 
  arrange(p_val) %>% 
  filter(matchStatus=='exactMatch') %>% 
  mutate(`Reference_PFAM` = Foreground/allForeground,
         `Query_PFAM` = BACKGROUND/allBackground) %>% 
  left_join(PDB_pfam %>% select(PFAMpref,PFAM_Name,PFAM_desc), by=c("PFAM" = "PFAMpref")) %>% 
  distinct() %>% 
  dplyr::select(Foreground,BACKGROUND, PFAM,PFAM_desc,Query_PFAM,Reference_PFAM) %>% 
  filter(Foreground >= 7) %>% select(-c(Foreground,BACKGROUND)) %>% 
  gather(key,value,-c(PFAM,PFAM_desc)) %>% 
  ggplot(aes(x=reorder(PFAM_desc,value,FUN=max),y=value,fill=key)) + 
  geom_bar(stat="identity",position="dodge", show.legend = TRUE) +
  coord_flip() +
  xlab("") + ylab("Proportion of PFAM terms") +
  labs(fill="")

ggsave("../results/F2b.pdf",width=8,height=3.5)
```



##Check putative additional domains in NEK kinases
EF hand domains in NEK kinases
```{r}
metrics_nMatch %>% filter(n_Match > 0, str_detect(Description,"NEK")) %>% 
  filter(nPFAM_PDB > n_Match) %>% 
  filter(code=="3HX4") %>% 
  left_join(metrics_long,by="GeneID") %>% 
  select(GeneID, n_Match, contains("PFAMpref"), contains("PFAM_desc")) %>% 
  distinct() %>% 
  filter(PDB_PFAMpref=="PF13499") 

```


Kinase domains hypothetical proteins
```{r}
metrics_nMatch %>% 
  filter(descBin=="Hyp") %>% 
  filter(nPFAM_PDB > n_Match) %>% 
  filter(code=="4O1O") %>% 
  left_join(metrics_long,by="GeneID") %>% 
  group_by(GeneID) %>% filter(!str_detect(GDB_PFAM_desc,'kinase')) %>% 
  select(GeneID, n_Match, contains("PFAMpref"), contains("PFAM_desc")) %>% 
  distinct() %>% count(GeneID)
         
```



##Figure 2c
Plot n. unique PFAM codes available via query and reference peptides.
```{r}
metrics_nMatch %>% select(GeneID, contains('nPFAM'),  n_Match, matchStatus) %>% 
  filter(matchStatus=="exactMatch") %>% 
  select(contains('nPFAM')) %>% 
  gather(key=source,value=nPFAM) %>% 
  ggplot(aes(x=nPFAM)) + geom_bar(aes(fill=source), position="dodge") +
  xlab("N. unique PFAM codes") + ylab("N. proteins")


metrics_nMatch %>% select(GeneID, contains('nPFAM'),  n_Match, matchStatus) %>% 
  filter(matchStatus=="exactMatch") %>% 
  select(contains('nPFAM')) %>% 
  gather(key=source,value=nPFAM) %>% 
  group_by(source) %>% summarize(mean_nPFAM=mean(nPFAM)) 

ggsave("../results/F2c.pdf",width=5,height=4)

```





#Random forest classifier

##Train classifier
Create training and test data
```{r, eval=TRUE}

Gd_forRF <- metrics_nMatch %>% select(1,TM:RMSD_model_sd,lenRatio,SS_sd,matchStatus) %>% na.omit()
#Gd_forRF %>% str()
  
geneOrder <- read_tsv("../data/Gd_forRF_geneIDorder.tsv")

Gd_forRF <- geneOrder %>% left_join(Gd_forRF, by="GeneID")  #original geneOrder


set.seed(1234) ; forTRAIN_exact <- Gd_forRF %>% filter(matchStatus=="exactMatch") %>%  sample_n(750);
set.seed(1234) ; forTRAIN_noMatch <- Gd_forRF %>% filter(matchStatus!="exactMatch") %>%  sample_n(750);
forTRAIN <- rbind(forTRAIN_exact, forTRAIN_noMatch)

#set.seed(1234); TEST <- Gd_forRF %>% anti_join(forTRAIN, by="GeneID") %>% sample_n(1500) ; 

set.seed(4321); forTEST_exact <- Gd_forRF %>% anti_join(forTRAIN_exact, by="GeneID") %>% filter(matchStatus=="exactMatch") %>% sample_n(250)
      
set.seed(4321); forTEST_noMatch <- Gd_forRF %>% anti_join(forTRAIN_noMatch, by="GeneID") %>% filter(matchStatus!="exactMatch") %>% sample_n(250)

TEST <- rbind(forTEST_exact, forTEST_noMatch) 

TRAIN <- forTRAIN %>% dplyr::select(-c(GeneID)) 
```

Model training is only run once.
```{r, eval = FALSE}
rf_model <- train(matchStatus ~. , data=TRAIN, method="rf",
                  trControl = trainControl(method="cv", number=5, verboseIter=TRUE),
                  prox=TRUE, verbose=TRUE)
saveRDS(rf_model,"../data/rf_model.Rds")
```

Print final model
```{r}
rf_model <- readRDS("../data/rf_model.Rds")

print(rf_model$finalModel)

```



##Classifier performance
```{r}

predTEST <- predict(rf_model, newdata=TEST, type="prob")
predALL <- predict(rf_model, newdata=Gd_forRF, type="prob")

#summary(predTEST)
#summary(predALL)


#accuracy on hold-out TEST set:
TEST %>%  mutate(RFexact_pred = predTEST$exactMatch) %>% 
  mutate(predStatus=ifelse(RFexact_pred >= 0.5,"exactMatch","noMatch")) %>% 
  count(matchStatus, predStatus)

#entire proteome
Gd_forRF %>% mutate(RFexact_pred = predALL$exactMatch) %>% 
  mutate(predStatus=ifelse(RFexact_pred >= 0.5,"exactMatch","noMatch")) %>% 
  count(matchStatus, predStatus)

Gd_matchPredict <- Gd_forRF %>% mutate(exactMatchPred = predALL$exactMatch) %>% ungroup()


```

##Fig 3a
###Compute feature importance 
```{r}

Gd_imp <- data.frame(varImp(rf_model, scale=FALSE)$importance)  

Gd_imp$noi  <- row.names(Gd_imp); row.names(Gd_imp) <- NULL
names(Gd_imp) <- c('Contribn', 'impFactor') ; #head(Gd_imp)
Gd_imp <- Gd_imp %>%dplyr::select(2,1) %>% arrange(desc(Contribn)) %>%  mutate(propCont = Contribn/sum(Contribn))

imp_plot <- Gd_imp %>% 
  mutate(impFactor_recode = case_when(impFactor=="AApc" ~ 'AA_%ID',
                               impFactor=="Cscore" ~ 'C_score',
                               impFactor=="AApc" ~ "AA_%ID",
                               impFactor=="Cov" ~ 'Coverage',
                               impFactor=="exactMatchPred" ~ 'Exact_match_prediction',
                               impFactor=="lenRatio" ~ 'Length_ratio',
                               TRUE ~ impFactor)) %>% 
  ggplot(aes(x=reorder(impFactor_recode, propCont), y=propCont)) +
  geom_bar(stat="identity", aes(fill=impFactor_recode), show.legend = FALSE) +
  ylab("Importance") + xlab("") + ylim(0,0.3) +
  coord_flip()  

imp_plot

ggsave("../results/SF3a.pdf",imp_plot, width=4,height=3)

```


##Fig 3b
###Sensitivity and specificity

```{r}

##AUROC

#https://gist.github.com/jwaage/6d8f4eb096e4f18a0894ca1ce27af834

forROC <- Gd_matchPredict %>% 
  dplyr::select(TM:exactMatchPred) %>% 
  mutate(matchStatus = abs(as.numeric(factor(matchStatus))-2)) 

aucOut <- data_frame('sensPlt' = NA, 'specPlt' = NA, 'Metric'=NA,'AUC' = NA)
for (i in names(forROC)){#print(i)
  r <- roc(forROC$matchStatus, forROC[ , i] %>% pull())
  sens <- r$sensitivities
  spec <- r$specificities
  a <- auc(r)[1]
  rnd <- round(a,3)
  result <- data_frame('sensPlt' = rev(sens), 'specPlt' = rev(spec)) %>% mutate(Metric = i, AUC = rnd)
  aucOut <- rbind(aucOut,result) %>% na.omit() }

auc_plot <- aucOut %>%  mutate(Metric = case_when(Metric=="AApc" ~ 'AA_%ID',
                               Metric=="Cscore" ~ 'C_score',
                               Metric=="AApc" ~ "AA_%ID",
                               Metric=="Cov" ~ 'Coverage',
                               Metric=="exactMatchPred" ~ 'Exact_match_prediction',
                               Metric=="lenRatio" ~ 'AA_length_ratio',
                               TRUE ~ Metric)) %>% 
  filter(Metric !="matchStatus") %>% 
  distinct() %>% na.omit() %>% filter(AUC >= 0.7) %>% 
  ggplot(aes(x=specPlt,y=sensPlt)) + 
  geom_segment(aes(x = 0, y = 1, xend = 1, yend = 0), alpha = 0.5) + 
  geom_step(aes(col=Metric), lty = 2,lwd=1) +
  scale_x_reverse(name = "False positive rate (Specificity)",limits = c(1,0)) + 
  ylab("True positive rate (Sensitivity)") + 
  coord_equal() +
  scale_color_manual(values= c("AA_%ID" = "#F8766D","C_score" = "#DE8C00",
                               "Exact_match_prediction" = 'black','AA_length_ratio' = "#00BA38",
                               "RMSD" = "#00BFC4","TM_model" = "#F564E3")) 

aucOut %>% select(-contains('Plt')) %>% distinct() %>% arrange(desc(AUC))

auc_plot

ggsave("../results/F3b.pdf",auc_plot, width=5,height=5)

```


##Test classifier on models from Human
```{r}

Hs_TEST <- read_tsv("../data/Hsapiens_TESTdata.tsv", col_types = cols())

write_tsv(Hs_TEST, path="../results/STable1.tsv")

#confirm that Gd TRAINing set and Hs_TEST set have same format
#TRAIN %>% str()
#Hs_TEST[,3:14] %>% str()

pred_Hs <- as_tibble(predict(rf_model, newdata=Hs_TEST[,3:14], type="prob"))
Hs_TESTpred <- cbind(Hs_TEST,pred_Hs) 

Hs_TESTpred %>% mutate(pred_matchStatus = ifelse(exactMatch > 0.5,'exact','noMatch')) %>% 
  count(matchStatus, pred_matchStatus) %>% 
  group_by(matchStatus) %>% mutate(gpErr=n/sum(n))

```


##Metric distribution by RF confidence
```{r}

Gd_all <- metrics_nMatch %>% left_join(Gd_matchPredict %>% dplyr::select(1,exactMatchPred),by="GeneID")  

Gd_all_RFconf <- Gd_all %>% 
  mutate(RF_confidence = case_when(matchStatus=="exactMatch" & exactMatchPred >= 0.5 ~ 'HiConf',
                                matchStatus=="noMatch" & exactMatchPred < 0.5        ~ 'LowerConf',
                                matchStatus=="exactMatch" & exactMatchPred < 0.5     ~ "LowerConf-like",
                                matchStatus=="noMatch" & exactMatchPred >= 0.5       ~ "HiConf-like",
                                TRUE ~ "")) 


Gd_all_RFconf %>% select(TM:RMSD_model_sd,lenRatio,SS_sd,matchStatus,exactMatchPred,RF_confidence) %>% 
  rename(Exact_match_prediction = exactMatchPred, AA_length_ratio = lenRatio,
         `AA_%ID` = AApc, Coverage = Cov, C_score = Cscore) %>% 
  filter(C_score > -8 & RMSD_model < 25) %>% 
  gather(key,value,-c(RF_confidence,matchStatus)) %>% 
  filter(str_detect(RF_confidence,"Conf")) %>% 
  na.omit() %>% 
  mutate(key = fct_relevel(factor(key), c('SS_sd', 'C_score','C_sd', 'AA_length_ratio',
                                          'TM','RMSD','AA_%ID','Coverage',
                                          'TM_model','TM_model_sd','RMSD_model','RMSD_model_sd',
                                          'Exact_match_prediction'))) %>% 
  ggplot(aes(x=key, y=value, col=RF_confidence)) + 
  geom_point(position=position_jitterdodge(jitter.width = 0.1, dodge.width=1), 
             size=0.02, alpha=0.2, show.legend = TRUE) +
  geom_boxplot( position=position_dodge(1), alpha=0) +
  xlab("") + ylab("") +
  theme(legend.title=element_blank(), 
        strip.background = element_blank(), 
        strip.text.x = element_text(size=0)) +
  facet_wrap( ~ key, scales="free")

ggsave("../results/F4.pdf",width=9.5,height=9)

```




##SFigure 4a
Technical variation: Train 500 models on the same data.
```{r, eval = FALSE}

RF_train_iter500 <- data_frame('dummy'=rep(NA, nrow(Gd_forRF)))
for(i in 1:500){
  
  a <- train(matchStatus ~. , data=TRAIN, method="rf",
             trControl = trainControl(method="cv", number=5, verboseIter=TRUE),
             prox=TRUE, verbose=TRUE)
  
  fm <- as_data_frame(predict(a, newdata = Gd_forRF, type="prob"))
  fm <- fm[ , 1]
  names(fm) <- paste0('rf_',i)
  RF_train_iter500 <- cbind(RF_train_iter500, fm)
  
}

RF_train_iter500 <- RF_train_iter500[ , -1]

saveRDS(RF_train_iter500, "../data/RF_train_iter500.Rds")
```
 
Plot technical variation
```{r}

RF_train_iter500 <- readRDS("../data/RF_train_iter500.Rds")

RF_train_iter500_summary <- data_frame('mean' = rowMeans(RF_train_iter500), 
                      'sd' = apply(RF_train_iter500 ,1, sd, na.rm = TRUE),
                      'sem' = apply(RF_train_iter500 , 1, function(x) sd(x)/sqrt(length(x))),
                      'len' = apply(RF_train_iter500 , 1, length ))
RF_train_iter500_summary %>% cbind(Gd_forRF) %>% dplyr::select(mean,sd,sem,matchStatus) %>% 
  ggplot(aes(mean, log10(1/sd))) + geom_point(aes(col=matchStatus), cex=0.5) +
  ylim(1,3.5)

ggsave("../results/SF6a.pdf",width=5,height=4)

```





##SFigure 4b

Robustness: Train 50 models on different training data and set sizes
```{r, eval=FALSE}
RF_sample_iter500 <- data_frame("GeneID" = Gd_forRF %>% select(GeneID) %>% pull())

for(i in c(50,100,200,300,400,500)){ 
  for(j in 1:50){
    sample_iterTrain <- rbind(Gd_forRF %>% filter(matchStatus=='noMatch') %>% sample_n(i),
                       Gd_forRF %>% filter(matchStatus=='exactMatch') %>% sample_n(i))
    
    rfMod_iter <- train(matchStatus ~ . , data = sample_iterTrain %>% select(-GeneID), method="rf",
                        trControl = trainControl(method="cv", number=5, verboseIter=TRUE),
                        prox=TRUE, verbose=TRUE)
    fm_t <- as_data_frame(predict(rfMod_iter, newdata=Gd_forRF, type="prob"))
    fm_t <- fm_t[ , 1]
    names(fm_t) <- paste('rf_train', 2*i, 'iter', j, sep="_")
    RF_sample_iter500 <- cbind(RF_sample_iter500, fm_t)
  }
}

saveRDS(RF_sample_iter500, "../data/RF_sample_iter500.Rds")

```

Plot reproducibility
```{r, eval=TRUE}

RF_sample_iter500 <- readRDS("../data/RF_sample_iter500.Rds")

RF_sample_iter500 %>% 
  gather(key,value,-GeneID) %>% 
  separate(key,into=c("v1","v2","v3","v4","v5"),sep="_",convert = TRUE) %>% 
  select(GeneID,v3,v5,value) %>% dplyr::rename(trainingSetSize=v3,iteration=v5) %>%
  group_by(GeneID,trainingSetSize) %>% summarize(mean=mean(value),sd=sd(value)) %>% 
  left_join(Gd_forRF %>% select(GeneID,matchStatus),by="GeneID") %>% 
  ggplot(aes(x=factor(trainingSetSize),y=sd)) +
  geom_boxplot(alpha=0.2, aes(col=matchStatus )) +
  xlab('Training set size') + ylab("sd(Probability of high-confidence category)")

ggsave("../results/SF6b.pdf",width=7,height=4)
```


##SFigure 4c
```{r}
RF_sample_iter500 %>% 
  gather(key,value,-GeneID) %>% 
  separate(key,into=c("v1","v2","v3","v4","v5"),sep="_",convert = TRUE) %>% 
  select(GeneID,v3,v5,value) %>% 
  dplyr::rename(trainingSetSize=v3,iteration=v5) %>%
  group_by(GeneID,trainingSetSize) %>% 
  mutate(call=ifelse(value>0.5,1,0)) %>% 
  group_by(GeneID,trainingSetSize) %>% 
  mutate(meanCall=mean(call), meanVal=mean(value),
         sdCall=sd(call), sdVal=sd(value)) %>% 
  ungroup() %>% 
  select(-c(value,iteration,call)) %>% distinct() %>% 
  left_join(Gd_forRF %>% select(GeneID, matchStatus),by="GeneID") %>% 
  left_join(Gd_all_RFconf %>% select(GeneID,exactMatchPred,RF_confidence),
            by="GeneID") %>% filter(RF_confidence!="") %>% 
  ggplot(aes(x=meanCall,y=meanVal)) +         
  geom_point(aes(col=RF_confidence),cex=0.2) + 
  facet_wrap(~ trainingSetSize) 

ggsave("../results/SF6c.pdf",width=8,height=4)
```



##Features of lower conf-like models?
```{r}
Gd_all_RFconf %>% filter(RF_confidence=="LowerConf-like") %>% 
  left_join(metrics_long,by="GeneID") %>% 
  filter(PDB_PFAMpref== GDB_PFAMpref) %>% 
  select(GeneID,GDB_PFAMpref) %>% 
  distinct() %>% 
  left_join(PFAM_descriptionTable,by=c('GDB_PFAMpref' = 'PFAMcode')) %>% 
  count(desc) %>% arrange(desc(n)) %>% filter(n>1)

Gd_all_RFconf %>% filter(RF_confidence=="LowerConf") %>% 
  left_join(metrics_long,by="GeneID") %>% 
  select(GeneID,GDB_PFAMpref) %>% 
  na.omit() %>% 
  distinct() %>% 
  left_join(PFAM_descriptionTable,by=c('GDB_PFAMpref' = 'PFAMcode')) %>% 
  count(desc) %>% arrange(desc(n)) %>% filter(n>5)



```


##Compute relative abundance of domains across confidence groups
```{r}

Gd_all_RFconf %>% 
  left_join(metrics_long,by="GeneID") %>% 
  select(GeneID,GDB_PFAMpref,PDB_PFAMpref, RF_confidence) %>%distinct() %>%  
  gather(key,value,-c(GeneID,RF_confidence)) %>% 
  distinct() %>% ungroup() %>% 
  filter(RF_confidence!="") %>% na.omit() %>% 
  group_by(RF_confidence,key) %>% count(value) %>% 
  spread(RF_confidence,n, fill = 0) %>% 
  filter(key=="GDB_PFAMpref") %>% 
  rowwise() %>% 
  mutate(ratio = (LowerConf /1389) / (sum(`HiConf`,`HiConf-like`,`LowerConf-like`) / 1389)) %>% 
  filter(LowerConf>0, ratio!="Inf") %>% 
  arrange(desc(ratio)) %>% 
  left_join(PFAM_descriptionTable,by=c('value'='PFAMcode')) %>% 
  select(desc,ratio, everything())
  
  
    
```




#Clustering analysis


##SFigure 5
Compare sequence- and structure-based clusters
```{r clustering, eval=TRUE}

blst_Nek <- read_tsv("../data/Nek_kinase_BLASTp.tsv", col_names = TRUE)
TM_Nek <- read_tsv("../data/Nek_kinase_TM.tsv", col_names = TRUE)

blst_matFull <- blst_Nek %>% mutate(value=percent_rank(bit_score)) %>% 
  select(query_id,subject_id,value) %>% 
  spread(subject_id,value, fill = 0) %>% select(-query_id) %>% as.matrix() 
rownames(blst_matFull) <- colnames(blst_matFull)

TM_matFull <- TM_Nek %>% spread(struc2,TM, fill=0) %>% select(-struc1) %>% as.matrix()
rownames(TM_matFull) <- colnames(TM_matFull)


blst_TM_mds <- rbind((1-cor(blst_matFull)) %>% 
                       cmdscale() %>%
                       as_tibble() %>% 
                       mutate(dtype="blst"), 
                     (1-cor(TM_matFull)) %>% 
                       cmdscale() %>%
                       as_tibble() %>% 
                       mutate(dtype="TM")) %>% 
  dplyr::rename(Dim.1 = V1, Dim.2 = V2) %>% 
  mutate(GeneID=str_replace(rep(rownames(blst_matFull),2),"GL_","GL50803_"))


blst_TM_mds %>% 
  left_join(Gd_all_RFconf, by="GeneID") %>% 
  mutate(dtype=recode(dtype,blst="BLAST",TM="TM-align")) %>% 
  filter(!is.na(RF_confidence)) %>%
  dplyr::rename("RF confidence" = RF_confidence) %>% 
  ggplot(aes(x=Dim.1,y=Dim.2 )) + 
  geom_point(aes(col=`RF confidence`, shape=`RF confidence`), size=2) +
  scale_shape_manual(values=c('HiConf'=1,'HiConf-like' = 1, 'LowerConf' =2, "LowerConf-like" = 2)) +
  facet_wrap(~ dtype , scales="free")

ggsave("../results/SF7.pdf",width=9,height=4)

```



#Transcriptional abundance
##SFigure 6
Transcriptional abundance by RF confidence
```{r, message='hide', warning='none'}

transcription_cpm <- read_tsv("../data/Ansell_2017_GiardiaCPM.tsv", col_types = cols())


transc_by_category <- Gd_all_RFconf %>% 
   mutate(transcriptLen=(GDB_AA * 3) - 3) %>% 
   left_join(transcription_cpm,by="GeneID") %>% 
   select(GeneID,RF_confidence,transcriptLen,contains('-s')) %>% 
   gather(key,value,-c(GeneID,transcriptLen,RF_confidence)) %>% 
   mutate(value=ifelse(is.na(value),0.1,value)) %>% 
   mutate(value=value/transcriptLen) %>% na.omit() %>% 
   group_by(GeneID,RF_confidence) %>% summarize(meanVal=mean(value)) %>% 
  filter(!RF_confidence=="") 
 
transc_by_category %>% 
   filter(meanVal > 0.001) %>% 
   ggplot(aes(x=RF_confidence,y=log10(meanVal))) + 
   geom_boxplot(aes(fill=RF_confidence), width=0.5, col="black", alpha=0.5) +
   geom_violin(aes(col=RF_confidence), alpha=0) +
   theme(legend.position="none") +
   theme(axis.text.x=element_text(angle=45, hjust=1)) + 
   xlab("") + ylab("log10(length-normalized CPM)")

ggsave("../results/SF8.pdf",width=4,height=4)

#Test differences 
aov_transc.res <- aov(meanVal ~ RF_confidence, 
                      data=transc_by_category %>% filter(meanVal> 0.001)) 
summary(aov_transc.res)  

TukeyHSD(aov_transc.res)
 
```

#Supplementary Tables

Write supplementary Tables
```{r}

write_tsv(Gd_all_RFconf %>% 
            select(-c(seq.ssLen,
                      nHrc,nErc,nCrc,
                      strucSum,strpPDB_AA, codechain)) %>% 
                     dplyr::rename(Exact_match_prediction =exactMatchPred,
                Length_ratio=lenRatio,Match_status=matchStatus,`AA_%ID`=AApc,
                C_score=Cscore,
                Annot_status=descBin,Confidence_category =RF_confidence) %>% 
            arrange(desc(C_score,Confidence_category)),
            path="../results/STable_2.tsv")

write_tsv(Gd_all_RFconf %>% 
            select(-c(seq.ssLen,
                      nHrc,nErc,nCrc,
                      strucSum,strpPDB_AA, codechain)) %>% 
            filter(RF_confidence=="HiConf-like" & descBin=="Hyp") %>%
            dplyr::rename(Exact_match_prediction =exactMatchPred,
                Length_ratio=lenRatio,Match_status=matchStatus,`AA_%ID`=AApc,
                C_score=Cscore,
                Annot_status=descBin,Confidence_category =RF_confidence) %>% 
              arrange(desc(C_score,Confidence_category)), 
          path="../results/STable_3.tsv")

```







#RF clasifier sans %AA ID
##Figure 3b
RF classifier without AA_%ID
```{r, eval=FALSE}

TRAIN_nopcAA <- forTRAIN %>% dplyr::select(-c(GeneID,AApc)) 

rf_model_nopcAA <- train(matchStatus ~. , data=TRAIN_nopcAA, method="rf",
                  trControl = trainControl(method="cv", number=5, verboseIter=FALSE),
                  prox=TRUE, verbose=FALSE)
saveRDS(rf_model_nopcAA, "../data/rf_model_nopcAA.Rds")
```

```{r}

rf_model_nopcAA <- readRDS("../data/rf_model_nopcAA.Rds")
print(rf_model_nopcAA$finalModel)

```



Classifier performance without % AA ID.
```{r}

predTEST_nopcAA <- predict(rf_model_nopcAA, newdata=TEST, type="prob")
predALL_nopcAA <- predict(rf_model_nopcAA, newdata=Gd_forRF, type="prob")

# summary(predTEST_nopcAA)
# summary(predALL_nopcAA)

#accuracy on hold-out TEST set:
TEST %>%  mutate(RFexact_pred_nopcAA = predTEST_nopcAA$exactMatch) %>% 
  mutate(predStatus=ifelse(RFexact_pred_nopcAA > 0.5,"exactMatch","noMatch")) %>% 
  count(matchStatus, predStatus) %>% 
  group_by(matchStatus) %>% mutate(gpErr=n/sum(n))

#entire proteome
Gd_forRF %>% mutate(RFexact_pred_nopcAA = predALL_nopcAA$exactMatch) %>% 
  mutate(predStatus=ifelse(RFexact_pred_nopcAA > 0.5,"exactMatch","noMatch")) %>% 
  count(matchStatus, predStatus) %>% 
  group_by(matchStatus) %>% mutate(gpErr=n/sum(n))


```


Compute feature importance and predictive power
```{r}

Gd_imp_nopcAA <- data.frame(varImp(rf_model_nopcAA, scale=FALSE)$importance) 

Gd_imp_nopcAA$noi  <- row.names(Gd_imp_nopcAA); row.names(Gd_imp_nopcAA) <- NULL
names(Gd_imp_nopcAA) <- c('Contribn', 'impFactor') 
Gd_imp_nopcAA <- Gd_imp_nopcAA %>%dplyr::select(2,1) %>% 
  arrange(desc(Contribn)) %>%  mutate(propCont = Contribn/sum(Contribn))

imp_plot_nopcAA <- Gd_imp_nopcAA %>% 
  mutate(impFactor_recode = case_when(impFactor=="Cscore" ~ 'C_score',
                               impFactor=="Cov" ~ 'Coverage',
                               impFactor=="exactMatchPred" ~ 'Exact_match_prediction',
                               impFactor=="lenRatio" ~ 'Length_ratio',
                               TRUE ~ impFactor)) %>% 
  ggplot(aes(x=reorder(impFactor_recode, propCont), y=propCont)) +
  geom_bar(stat="identity", aes(fill=impFactor_recode), show.legend = FALSE) +
  ylab("Importance") + xlab("") +
  coord_flip()  

imp_plot_nopcAA

ggsave("../results/SF5.pdf", imp_plot_nopcAA, width=4,height=3)
```
